Eco-evolutionary dynamics of finite populations from first principles

### 2023-02-27 ind_files/03_BD_processes.tex
Add comment on Ito formula

or, in integral form,

@@

Naively, we may expect the same logic to still hold true for one-dimensional It\^o processes of the form \eqref{ito_SDE_diff} with $gdt$ simply being replaced by $GdB_t$ on the RHS. However, this intuition is not true. The correct relation is given by 
\emph{It\^o's formula}\footnote{It\^o's formula also additionally requires $h \in C^2(\mathbb{R})$, meaning that $h$ is continuous and the first and second derivatives of $h$ exist and are also continuous}, which states:

@@

where the presence of the stochastic term has added an entire extra term $G^2(X_t)h''(X_t)/2$ that does not exist in the deterministic setting, in some sense because the Brownian motion `wiggles too erratically' over time. Using It\^o's formula and some simple algebra, one can then show that if a process $X_t$ taking values in $\mathbb{R}$ satisfies the SDE \eqref{ito_SDE_diff}, then the probability density $P(x,t)$ of finding the process in a state $x \in \mathbb{R}$ at time $t$ (if it exists) satisfies the PDE

>>>>

\emph{i.e.}

@@

Naively, we may expect the same logic to still hold true for one-dimensional It\^o processes of the form \eqref{ito_SDE_diff} with $gdt$ simply being replaced by $GdB_t$ on the RHS. However, this does not work. The correct relation is instead given by 
\emph{It\^o's formula}\footnote{It\^o's formula also additionally requires $h \in C^2(\mathbb{R})$, meaning that $h$ is continuous and the first and second derivatives of $h$ exist and are also continuous}:

@@

Note that there is now an extra $h''(X_t)G^2(X_t)/2$ term that does not exist in the deterministic setting. In some sense, this term is present because the random fluctuations of Brownian motion are `too erratic' and do not follow our deterministic intuitions. Using It\^o's formula and some simple algebra, one can then show that given any process $X_t$ taking values in $\mathbb{R}$ satisfying the SDE \eqref{ito_SDE_diff}, the associated probability density $P(x,t)$ of finding the process in a state $x \in \mathbb{R}$ satisfies the PDE
